The fundamental shape of dynamic experiences is preserved in episodic memory

### 2018-07-09 main.tex
checking in intro...struggling on the framing

Another major difference between traditional trial-based memory paradigms and real-world memory concerns the subjective experience of the participant.  For example, consider an emotionally salient event in your life (e.g.\ marriage, the birth of a child, death of a loved one, etc.), or even a movie that you were especially impacted by.  No list-learning paradigm (or similar) can hope to capture the nuance and depth of such experiences.  Our everyday experiences feel ``important'' to us, whereas the words we study on a random word list do not.  One component of naturalistic experiences that enhances their impact concerns the temporal structure with which they unfold.  For example, our experiences in the real world are necessarily autocorrelated in space and time on short timescales.  Further, or experiences are often characterized by longer timescale correlations that reflect the impact of our past actions and observations.  These temporal dynamics are not typically present in traditional memory studies, but are important if we wish to understand how our memory systems remember our everyday experiences.

Despite the fact that information flows continuously through our sensory systems, prior research suggests that our memories are partitioned and centered around particular ``events'' \citep{Radv12, RadvZack11, BrunEtal18}. Short timescale autocorrelations in the contents of our experiences may cause information to become associatively linked in memory, while sharp contrasts (i.e. ``event boundaries'') may segment continuous experiences into events \citep{HeusEtal18, BrunEtal18, Radv12}. There is compelling evidence that event boundaries can influence memory organization in trial-based recognition \citep{HeusEtal18}, navigation \citep{BrunEtal18}, and narrative comprehension \citep{ZwaaRadv98, EzzyDava11}.  However, the extent to which people naturally (verbally) recount their experiences in segments centered around particular events has not been explored.

To study the temporal and semantic aspects of memory for naturalistic experiences, we analyzed an open dataset in which participants view and verbally recall an episode of the BBC series \textit{Sherlock}  \citep{ChenEtal17}. We developed a novel computational approach based on topic modeling \citep{BleiEtal03} and hidden Markov models \citep{Rabi89, BaldEtal17} to capture the temporal evolution of rich semantic structure present in naturalistic experiences. We then use this model to characterize verbally recalled episodic memories for naturalistic experiences. We hypothesized that memories for naturalistic experiences are 1) segmented and centered around particular events and 2) less about recalling specific details and more about preserving the overall ``shape'' of an experience.

>>>>

A defining feature of naturalistic experiences is the temporal structure with which they unfold. Our experiences in the real world are typically autocorrelated in space and time on short timescales. In other words, the contents of an experience are often similar from moment to moment and change gradually. For example, many of us spend our mornings/evenings at home and much of our days at work (e.g. an office). While the specific happenings within these contexts differs, they are (usually) highly correlated and predictable over time. Furthermore, our experiences can also be characterized by longer timescale correlations. For example, consider returning to your office after taking a lunch break. The perceptual contexts before and after lunch are highly similar, but separated in time. However, these recurrent and gradually changing temporal dynamics are not typically present in traditional memory studies, but are likely important if we wish to understand how our memory systems remember our everyday experiences.

Despite the continuous and (often) gradually changing nature of naturalistic experiences, prior research suggests that our memories for those same experiences may be more descrete and centered around particular ``events'' \citep{Radv12, RadvZack11, BrunEtal18}. Short timescale autocorrelations in the contents of an experience may cause information to become associatively linked in memory (and thus treated as a single event), while sharp changes (i.e. ``event boundaries'') may segment continuous experiences into events \citep{HeusEtal18, BrunEtal18, Radv12}. There is now compelling evidence that event boundaries can influence memory organization when learning sequences of items \citep{HeusEtal18, DuBrDava13}, during navigation \citep{BrunEtal18}, and narrative comprehension \citep{ZwaaRadv98, EzzyDava11}.  Furthermore, brain activity while viewing naturalistic stimuli displays a block-diagonal structure, and the neurally-derived segments are correlated with judgements of event boundaries \citep{BaldEtal17}.  However, whether people naturally recount their experiences in as events has not been explored.

% We propose that variability in the contents of an experience over time define a ``stimulus-space''. In this space, timepoints where the contents of experience are more similar are represented as points that are closer together. Timepoints with relatively more distinct content are represented further apart. The moment-by-moment flow of content through this space can be represented as a ``stimulus trajectory''.
%
% As a consequence of this structure, we hypothesized that verbally recounted memories for those experiences would mirror this shape.

To study the temporal and semantic aspects of memories for naturalistic experiences, we analyzed an open dataset in which participants view and then verbally recall an episode of the BBC series \textit{Sherlock}  \citep{ChenEtal17}. We Furthermore, we hypothesized that the stimulus To test these hypotheses, we developed a novel computational approach based on topic modeling \citep{BleiEtal03} and hidden Markov models \citep{Rabi89, BaldEtal17} to capture the temporal structure present in naturalistic experiences. We then use this model to characterize the shape of the stimulus as well as verbally recalled memories.
